Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-02-26 write-ups/FSE/paper.tex
Update figure overview
Pass over new section.

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{overview.pdf}
\caption{Cooperative kernels can flexibly resize to let run other tasks
(e.g. graphics) concurrently}
\label{fig:overview}
\end{figure}

@@

Some recent GPUs, including NVIDIA's Pascal architecture, provide
hardware support for instruction-level preemption.  Few details of how
this works are available, however from reading NVIDIA documentation
and online material~\cite{PascalWhitepaper,anandtech}, and running
litmus tests on a Pascal platform, our understanding is that
preemption of entire kernels, but not of individual workgroups running
a kenrel is supported.  Intel---the only vendor whose OpenCL
implemenation is capable enough to implement cooperative kernels---do
not provide this feature, thus we cannot make a direct comparison.  We
present here a theoretical analysis of the overheads associated with
sharing the GPU between graphics and compute tasks via kernel-level
preemption.

Let us assume that a graphics workload is required to be scheduled
with period $P$ and duration $D$, and that a long-running compute
kernel requires time $C$ to execute without interruption.  If we
assume the cost of preemption is negligible (e.g.\ NVIDIA have
reported preemption times of 0.1 ms for Pascal), then the overhead
associated with switching between compute and graphics every $P$ time
steps is $P/(P-D)$. We compare this task-level preemption overhead
model with our experimental results per graphics workload in
\mytab{\ref{tab:preemption}}.

We see that our approach provides lower overhead for low and medium
graphics workloads. Recall for the heavy graphics workload, we are
within $10\%$ of the required deadline ($40$ ms) by using half the
workgroups with an overhead of $1.XX$, which is less than the
kernel-level preemption overhead of $1.33$. \TSComment{need to get
  updated numbers}

The low overhead of cooperative kernels for low and medium graphics
workloads are due to the irregularity of the cooperative kernels and
the property that they do not scale linearly. Thus some resources can
be taken away during computation without proportionally reducing
execution time.  Our poor performance for the heavy workload is
because the graphics task requires all but one workgroup, which
cripples the cooperative kernel enough that $\offerfork$ calls are not
issued as frequently.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{overview.pdf}
\caption{Cooperative kernels can flexibly resize to let other tasks
(e.g. graphics) run concurrently}
\label{fig:overview}
\end{figure}

@@

Some recent GPUs, including Nvidia's Pascal architecture, provide
hardware support for instruction-level preemption.  Few details of how
this works are available, but from reading Nvidia documentation
and online material~\cite{PascalWhitepaper,anandtech}, and running
litmus tests on a Pascal platform, our understanding is that
preemption of entire kernels, but not of individual workgroups running
a kernel is supported.  Intel GPUs do
not provide this feature, and out OpenCL 2.0 prototype of cooperative kernels cannot run on Nvidia GPUs, making a direct comparison impossible.
We present here a theoretical analysis of the overheads associated with
sharing the GPU between graphics and compute tasks via kernel-level
preemption.

Let us assume that a graphics workload is required to be scheduled
with period $P$ and duration $D$, and that a long-running compute
kernel requires time $C$ to execute without interruption.  If we
assume the cost of preemption is negligible (e.g.\ Nvidia have
reported preemption times of 0.1 ms for Pascal~\cite{TODO}), then the overhead
associated with switching between compute and graphics every $P$ time
steps is $P/(P-D)$.

We compare this task-level preemption overhead
model with our experimental results per graphics workload in
\mytab{\ref{tab:preemption}}.
%
Based on the above assumptions, our approach provides lower overhead for low and medium
graphics workloads. Recall for the heavy graphics workload, we are
within $10\%$ of the required deadline ($40$ ms) by using half the
workgroups with an overhead of $1.XX$, which is less than the
kernel-level preemption overhead of $1.33$. \TSComment{need to get
  updated numbers} \ADComment{I don't quite understand the wording; sounds a bit like we still win with the heavy workload.}

The low overhead of cooperative kernels for low and medium graphics
workloads is due to sub-linear scalability of the irregular algorithms that they implement: some resources can
be taken away during computation without proportionally reducing
execution time.  Our poor performance for the heavy workload is
because the graphics task requires all but one workgroup, which
cripples the cooperative kernel enough that $\offerfork$ calls are not
issued as frequently.
